\chapter{Specification}
uhms, I don't think this section will have interesting diagrams.
should it contain the theory? DSL, XML (vocabulary, grammar, motivations), intercommunication with other systems (config files, structure.xml...)

relate to reengineering (R Pressmann): abstraction, procedures, etc.

Users: no users (well, a "virtual" user does exist). it's a meta-app that reads apps in xml and renders them in html. define input and output.

contracts: no contracts. TDD as opposed to design by contract??? check literature

Use Cases: one per directive, settings and entities?? users: the system.


Larman: system sequence diagram + contracts of system operations
meyer: design by contract

Design by Contract and Test Driven Development are not at all mutually exclusive.
TDD is not about testing!! is about executable specs
"The first idea is about test-driven development not being a testing technique;
automatic tests only become ”tests” after having implemented all the functionality
being test-
driven
. Before that there is nothing to actually
test
and those artifacts
rather serve as executable
specifications
but not tests"

up to date documentation: a common problem is that docs are hardly ever updated


---
Extreme Programming advocates test driven development where tests are used
to specify the behavior of a program before the program code is actually written.
Together with using the simplest design possible and intention revealing program
code, tests are additionally used as a documentation of the program. However,
tests are not sufficient to completely define the behavior of a program because
they are only able to test properties of a program by example and do not allow to
state general properties. The latter can be achieved using formal specifications,
e.g. using Meyer’s design by contract % http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.68.313&rep=rep1&type=pdf